\documentclass[nonacm,sigconf]{acmart}

\usepackage{cite}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{tagging}
\usepackage{xurl}  % aggressive line-breaking for long \url{...} links
\usepackage{enumitem}
\usepackage{listings}
\usepackage{microtype}  % character protrusion / expansion for narrow columns

\lstset{
  basicstyle=\footnotesize\ttfamily,
  breaklines=true,
  breakatwhitespace=false,
  columns=flexible,
  frame=none,
  xleftmargin=0pt,
  xrightmargin=0pt,
  linewidth=\columnwidth,
  aboveskip=4pt,
  belowskip=4pt,
}

\usepackage{sc26repro}
\pagenumbering{gobble}
% Allow slightly looser word spacing
\emergencystretch=2em

% Comment out both lines before final submission.
%\usetag{explanation}
%\usetag{example}

\begin{document}

% Suppress acmart's running header (which references \shorttitle)
% since this appendix has no title block.
\pagestyle{empty}

\twocolumn[%
{\begin{center}
\Huge
Appendix: Artifact Description
\end{center}}
]

\appendixAD

\section{Overview of Contributions and Artifacts}

\subsection{Paper's Main Contributions}

\begin{description}[leftmargin=1.8em,labelindent=0pt,itemsep=3pt,topsep=3pt]
\item[$C_1$] \textbf{Structure-exploiting gradient computation for INLA.}
A custom differentiation rule for the full objective function,
covering factorization, and solve of BTA
precision matrices, that computes exact gradients while exploiting
block sparsity. By fusing stages of the forward computation, only
compact intermediate matrices need to be stored, reducing memory
by approximately $2\times$.
\item[$C_2$] \textbf{Extension to multivariate models.}
An extension of the gradient computation to jointly model $k$
correlated variables by exploiting the block structure of the
multivariate precision matrix, separating per-variable and
cross-variable hyperparameter contributions to reduce per-step
work (benchmarked up to $k{=}3$, $d{=}15$).
\item[$C_3$] \textbf{Multi-GPU distributed AD.}
For models whose stored intermediates exceed single-GPU memory,
we distribute the backward pass across multiple GPUs via
temporal-domain decomposition with nested dissection reordering.
We synchronize boundary intermediates via MPI allgather and stage
intermediates on CPU memory, enabling AD on models up to
1.9M latent variables that otherwise cannot run on a single GPU.
\item[$C_4$] \textbf{Comprehensive evaluation on up to 128 GH200 GPUs.}
Beyond reporting performance results, we provide a
practitioner-oriented analysis that disentangles algorithmic
gains from framework effects, evaluates when FD can match AD
via scaling, and characterizes where time and memory concentrate
in the gradient pipeline.
\end{description}

\subsection{Computational Artifacts}

Frameworks ($A_1$--$A_2$) are provided as Git repositories; every
per-figure benchmark ($A_3$--$A_{10}$) lives as a subdirectory inside
the single reproduction repository at\\
\url{https://github.com/affifboudaoud/adelia-artifact}\\
(commit \texttt{f2db10b}; Zenodo DOI to be assigned upon acceptance).

\begin{description}[leftmargin=2.5em,labelindent=0pt,itemsep=3pt,topsep=3pt]
\item[$A_1$] Serinv structured sparse solver:
  \url{https://github.com/affifboudaoud/serinv}
  (branch \texttt{serinv\_jax}, commit \texttt{5fa19bc}).
\item[$A_2$] DALIA/ADELIA framework:
  \url{https://github.com/affifboudaoud/DALIA}
  (branch \texttt{cleaning}, commit \texttt{9bdf6e6}).
\item[$A_3$] Gradient correctness suite
  (\texttt{validation/} subtree).
\item[$A_4$] Structure-preserving differentiation benchmark
  (\path{experiments/tab2_structure_comparison/}).
\item[$A_5$] L-BFGS convergence benchmark
  (AP1 + GST-T runs driven by \texttt{converge.sbatch}).
\item[$A_6$] End-to-end wall-clock benchmark
  (\path{experiments/fig2_wallclock/}).
\item[$A_7$] Problem-size scaling benchmark
  (\path{experiments/fig5_scaling_study/}).
\item[$A_8$] Framework-decomposition benchmark
  (\path{experiments/fig4_framework_decomposition/}).
\item[$A_9$] Resource and energy efficiency benchmark
  (\path{experiments/fig6_resource_efficiency/}).
\item[$A_{10}$] Per-stage distributed performance-analysis benchmark
  (\path{experiments/fig7_performance_analysis/}).
\end{description}

\begin{center}
\begin{tabular}{rll}
\toprule
Artifact ID  &  Contributions &  Related \\
             &  Supported     &  Paper Elements \\
\midrule
$A_1$   &  $C_1$         & --- (dependency) \\
$A_2$   &  $C_1$--$C_3$  & --- (framework) \\
\midrule
$A_3$   &  $C_1$--$C_3$  & correctness validation \\
$A_4$   &  $C_1$         & Table~2 \\
$A_5$   &  $C_1$--$C_3$  & Figure~3 \\
$A_6$   &  $C_1$--$C_3$  & Figure~4 \\
$A_7$   &  $C_1$, $C_4$  & Figure~5 \\
$A_8$   &  $C_4$         & Figure~6 \\
$A_9$   &  $C_3$, $C_4$  & Figure~7 \\
$A_{10}$ &  $C_3$, $C_4$ & Figure~8 \\
\bottomrule
\end{tabular}
\end{center}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Artifact Identification}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\newartifact

\artrel

$A_1$ is the Serinv structured sparse solver library. It provides
the block-tridiagonal (BT) and block-tridiagonal arrowhead (BTA) Cholesky factorization, triangular solve, and selected-inversion
algorithms on top of which the ADELIA custom backward pass is
implemented, relating to contributions $C_1, C_2, C_3$. $A_2$ depends on
$A_1$ for its solver infrastructure for the forward pass and for the finite-difference
baseline used throughout the evaluation.

\artexp

Working through this artifact results in the correct installation
of Serinv at the pinned commit and a successful import, so the downstream framework ($A_2$) and benchmark artifacts can load its algorithms. Serinv's own correctness is validated upstream; end-to-end numerical correctness of ADELIA's usage of Serinv is the subject of $A_3$.

\arttime

Setup and import check: $<$5~min.

\artin

\artinpart{Hardware}

CSCS Alps supercomputer, using NVIDIA GH200 Grace Hopper nodes.
Each node has four Hopper-based GPUs (96\,GiB HBM3 each) and a
72-core ARM Neoverse V2 (Grace) CPU with 480\,GiB LPDDR5X
memory, connected via NVLink-C2C. Multi-node runs communicate
over an HPE Slingshot-11 interconnect. A single GPU is
sufficient for this artifact.

\artinpart{Software}

Python~3.11, NumPy, SciPy, CuPy~13.6.0, CUDA~12. All
dependencies are pinned in the conda \texttt{environment.yml}
shipped with $A_2$.

\artinpart{Datasets / Inputs}

None.

\artinpart{Installation and Deployment}
To install the artifact (Serinv) and validate its correct installation, please execute the following:
\begin{lstlisting}[language=bash]
cd adelia-artifact
git submodule update --init serinv
pip install -e serinv/
python -c "import serinv; print(serinv.__file__)"
\end{lstlisting}

\artcomp

$T_1$: install the pinned Serinv package as an editable
dependency of $A_2$ (submodule of the reproduction repository).
$T_2$: verify the package imports cleanly.
Dependencies: $T_1 \rightarrow T_2$. No experimental parameters
or statistical repetition; the check is a deterministic
load-time verification.

\artout

The installed package loads without error, confirming that the
pinned Serinv commit is available to every downstream
benchmark artifact.


% ---- A2: DALIA/ADELIA framework ----
\newartifact

\artrel

$A_2$ is the DALIA/ADELIA framework. It wraps $A_1$ and implements
the JAX-based custom backward pass for the INLA objective, the
multivariate coregional extension, and the multi-GPU two-phase
distributed backward pass. It directly supports contributions
$C_1$, $C_2$, and $C_3$.

\artexp

Working through this artifact results in the correct installation
of ADELIA, verified by running the single-model gst\_small
gradient check from $A_3$ on one GPU. A successful run confirms
that the framework's AD entry points load correctly and produce
gradients matching the dense AD-Loop reference, which is the
precondition for all downstream benchmark artifacts
$A_4$--$A_{10}$.

\arttime

Installation: $\sim$20~min on CSCS Alps (including \texttt{mpi4jax}
compilation). Smoke test: $<$1~min.

\artin

\artinpart{Hardware}

Same as $A_1$.

\artinpart{Software}

In addition to the software listed for $A_1$:
JAX / JAXlib 0.8.1, \texttt{mpi4jax} 0.5.2+ (built from source
against Cray MPICH with CUDA, no CUDA-aware MPI or NCCL),
Pydantic $\geq$~2.0. Full version list and installation
walkthrough in \texttt{INSTALL.md}.

\artinpart{Datasets / Inputs}

None; functional verification uses synthetic inputs.

\artinpart{Installation and Deployment}

\begin{lstlisting}[language=bash]
conda env create -f environment.yml
conda activate adelia-env
pip install -e serinv/ -e DALIA/
export MPI4JAX_BUILD_MPICC=$(which mpicc)
pip install --no-binary mpi4jax "mpi4jax==0.8.1.post2"
\end{lstlisting}

\noindent
Due to a cuSolver/CuPy initialization conflict, JAX must be
initialized (including a GPU warmup) before any \texttt{from dalia
\ldots} imports happen. All experiment scripts handle this automatically.

\artcomp

$T_1$: install ADELIA and its runtime dependencies
(JAX/JAXlib, CuPy, mpi4jax).
$T_2$: run the smallest end-to-end gradient correctness check
from $A_3$ (on GST-S) to verify that ADELIA's AD entry points
load and produce gradients matching a dense reverse-mode JAX
reference.
Dependencies: $T_1 \rightarrow T_2$; $A_1 \rightarrow T_1$.
$T_2$ uses a single gradient evaluation (no statistical
repetition) because correctness is deterministic under fixed
random seeds.

\artout

A successful $T_2$ confirms that ADELIA is correctly installed
and that downstream benchmark artifacts $A_4$--$A_{10}$ can
proceed.  The framework has no quantitative outputs of its
own; its performance is evaluated through $A_4$--$A_{10}$.


% ---- A3: Gradient correctness suite ----
\newartifact

\artrel

$A_3$ is the gradient correctness suite (\texttt{validation/}). It
validates the exactness of ADELIA's custom backward pass by
comparing its gradients against finite differences (FD) and
against a dense reverse-mode JAX reference (AD-Loop) on all ten
benchmark models, covering contributions $C_1$, $C_2$, $C_3$.

\artexp

ADELIA's AD gradients should match AD-Loop within
$1.2 \times 10^{-7}$ relative error on models that fit in memory
(GST-S/M, GST-C2/C3) and match FD up to the FD truncation error
($O(h^2)$, $h{=}10^{-3}$) on all models. The distributed two-phase
path should match AD-Loop on GST-C3 with $P{=}2$ to within
$6.8 \times 10^{-6}$.

\arttime

Full sweep: $\sim$60~min on 1 GPU.
Quick single-model check (GST-S): $<$5~min.

\artin

\artinpart{Hardware, Software, Installation}

Same as $A_2$.

\artinpart{Datasets / Inputs}

Input matrices for the ten benchmark models are downloaded via
\path{scripts/download_data.sh} ($\sim$1.5\,GB total). Models
span univariate synthetic (GST-S/M/L), real-world temperature
(GST-T), multivariate synthetic (GST-C2/C3), and production-scale
multivariate (SA1, AP1, WA1, WA2) up to 1.9M latent variables
and $d{=}15$ hyperparameters.

\artcomp

The workflow runs once per model on all ten benchmark models.
$T_1$: load the model inputs and compute ADELIA's AD gradient
at a fixed hyperparameter point.
$T_2$: compute a reference gradient on the same point from
(a)~a dense reverse-mode JAX implementation (AD-Loop, where the
backward pass fits in GPU memory) and (b)~central finite
differences with step $h{=}10^{-3}$ on all models.
$T_3$: compare ADELIA against each available reference and
report forward-value relative error, gradient cosine similarity,
and max component-wise relative error.
Dependencies: $T_1, T_2 \rightarrow T_3$. No statistical
repetition: every check is deterministic under fixed seeds.

\artout

Per-model PASS/FAIL reports confirming gradient correctness.
Pass criteria are a forward-value relative error $< 10^{-6}$,
a gradient cosine similarity $> 0.999$, and a max
component-wise relative error $< 10^{-2}$ versus FD.
See \texttt{validation/README.md} for the full criteria list.


% ---- A4: Structure-preserving differentiation benchmark ----
\newartifact

\artrel

$A_4$ is the structure-preserving differentiation benchmark
(\path{experiments/tab2_structure_comparison/}). It times
five differentiation strategies (FD, AD-Dense, AD-Loop,
AD-Loop-Ckpt, AD-BTA) on five models of increasing size,
supporting contribution $C_1$ and producing Table~2.

\artexp

Only ADELIA's structure-preserving AD-BTA strategy should fit
all five models; AD-Dense, AD-Loop, and AD-Loop-Ckpt should
OOM on the larger models. AD-BTA should deliver the lowest
per-gradient time among AD strategies.

\arttime

$\sim$2~hours on a single GPU.

\artin

\artinpart{Hardware, Software, Installation, Inputs}

Same as $A_3$. Uses the GST-S, GST-C2, GST-C3, GST-M, GST-L
inputs already downloaded for $A_3$.

\artcomp

Five differentiation strategies (FD, AD-Dense, AD-Loop,
AD-Loop-Ckpt, AD-BTA) are exercised on five models of
increasing size (GST-S, GST-C2, GST-C3, GST-M, GST-L).
$T_1$: for each $(\text{model}, \text{strategy})$ pair, time
one gradient evaluation and record peak GPU memory;
$T_2$: aggregate per-pair results into Table~2, reporting
out-of-memory failures as OOM.
Dependencies: $T_1 \rightarrow T_2$. Each timed measurement
uses 2 warmup evaluations followed by 10 timed evaluations,
reporting the mean with 95\% confidence intervals. OOM
failures do not contribute to timing statistics.

\artout

Table~2 entries in the paper: per-model, per-strategy
per-gradient wall-clock time and peak GPU memory.  Aggregated
results live in \path{tab2_results.csv}; the plotting script
\path{plotting/scripts/plot_structure_comparison.py}
regenerates Table~2.


% ---- A5: L-BFGS convergence benchmark ----
\newartifact

\artrel

$A_5$ is the optimizer-convergence benchmark producing
Figure~3. It runs L-BFGS on the AP1 and GST-T models from the
same initialization using either ADELIA's exact gradients or
FD, and compares the objective and gradient-norm trajectories.
It supports contributions $C_1$--$C_3$ by showing that exact
gradients resolve FD stalling on real-world models.

\artexp

With exact gradients, L-BFGS should converge to a well-conditioned
minimum with gradient norm several orders of magnitude below the
FD baseline, which stalls.

\arttime

$\sim$8~hours combined across the two convergence runs
(AP1 on 4~GPUs, GST-T on 1~GPU).

\artin

\artinpart{Hardware, Software, Installation, Inputs}

Same as $A_3$. Uses the AP1 and GST-T datasets.

\artcomp

The workflow runs independently on AP1 and GST-T.
$T_1$: run L-BFGS from a fixed initial hyperparameter configuration
using ADELIA's exact gradients, logging the objective value
and gradient norm at every iteration.
$T_2$: repeat $T_1$ from the same initial point using the FD
gradient, on identical hardware.
$T_3$: align the $T_1$ and $T_2$ trajectories and produce the
convergence plot.
Dependencies: $T_1, T_2 \rightarrow T_3$. No statistical
repetition: L-BFGS is deterministic given the initial point.
Each run terminates on whichever comes first: the gradient-norm
tolerance, the maximum iteration count, or detection of an FD
stall.

\artout

Per-run L-BFGS logs with $f$ and $\|\nabla f\|$ per iteration
for each (model, method) pair. The plotting script
\path{plotting/plot_convergence_all.py} reads these logs and
produces Figure~3.


% ---- A6: End-to-end wall-clock benchmark ----
\newartifact

\artrel

$A_6$ is the end-to-end wall-clock benchmark
(\path{experiments/fig2_wallclock/}). It produces Figure~4 by
(1)~running each of the ten models to convergence under ADELIA
on $P_{\min}$ GPUs and recording both the per-iteration
wall-clock and the number of L-BFGS iterations to convergence,
and (2)~measuring the FD baseline's per-gradient wall-clock on
the same hardware. Figure~4 then reports each method's
end-to-end cost (optimization + Hessian) adjusted to the same
iteration count using ADELIA as the reference, isolating
computational cost from convergence differences. It supports
contributions $C_1$--$C_3$.

\artexp

When normalized to the same iteration count, ADELIA's
end-to-end wall-clock should be substantially lower than FD's
on every model, with $2.4$--$50.9\times$ per-gradient speedups
driving the gap, reproducing Figure~4.

\arttime

$\sim$8~hours total on 1--4 GPUs (small models together,
large models in separate jobs).

\artin

\artinpart{Hardware, Software, Installation, Inputs}

Same as $A_3$. Distributed models (SA1, AP1, WA1, WA2) run on
4 GPUs.

\artcomp

The workflow runs on all ten benchmark models at $P_{\min}$.
$T_1$: run ADELIA's L-BFGS optimizer to convergence on each
model, recording the per-iteration wall-clock and the number
of iterations $n_{\mathrm{iter}}$.
$T_2$: measure the FD baseline's per-gradient wall-clock on
each model using 2 warmup plus 10 timed evaluations with
95\% confidence intervals.
$T_3$: compute the Hessian cost for both methods (AD as two
backward passes per column; FD as $O(d^2)$ forward evaluations).
$T_4$: project both methods onto the same iteration count
$n_{\mathrm{iter}}$ and produce Figure~4, isolating
computational cost from convergence differences.
Dependencies: $T_1, T_2 \rightarrow T_3 \rightarrow T_4$.

\artout

Per-model timing logs aggregated into
\path{pipeline_wallclock.csv}. The plotting script
\path{plotting/scripts/plot_pipeline_wallclock_merged.py}
regenerates Figure~4.


% ---- A7: Scaling study ----
\newartifact

\artrel

$A_7$ is the problem-size scaling benchmark
(\path{experiments/fig5_scaling_study/}). It measures per-gradient
speedup of ADELIA over FD as a function of problem size on
single-GPU temporal variants of WA1 and spatial variants of WA2.
It supports $C_1$ (structured AD efficiency) and $C_4$ (scaling
characterization), producing Figure~5.

\artexp

Along both the temporal axis (WA1) and the spatial axis (WA2),
the AD-to-forward cost ratio $c_{\mathrm{AD}}$ should rise from
roughly $1.2$--$1.4$ at the smallest sizes to an asymptote near
$3.5$--$3.7$ as $\mathcal{O}(b^3)$ compute dominates
kernel-launch overhead, with per-gradient speedups over FD
correspondingly decreasing from about $25\times$ down to about
$8\times$ across the tested range, reproducing Figure~5.

\arttime

$\sim$6~hours on a single GPU (five temporal + three spatial
variants).

\artin

\artinpart{Hardware, Software, Installation}

Same as $A_3$.

\artinpart{Datasets / Inputs}

Temporal variants of WA1 at $n_s{=}1{,}247$
with $n_t \in \{2,\,32,\,64,\,128,\,256\}$,
and spatial variants of WA2 at $n_t{=}48$
with $n_s \in \{72,\,282,\,1{,}119\}$,
are pre-generated under \path{data/scaling_wa1_temporal/}
and \path{data/scaling_wa2_spatial/}.

\artcomp

The workflow runs independently on each problem-size variant
(five temporal variants of WA1, three spatial variants of WA2),
all on a single GPU so that distribution overhead does not
confound the scaling trend.
$T_1$: time one ADELIA forward pass, one ADELIA gradient, and
one FD evaluation on each variant using 2 warmup plus 10 timed
evaluations with 95\% confidence intervals.
$T_2$: compute the AD cost ratio
$c_{\mathrm{AD}} = t_{\mathrm{AD}} / t_{\mathrm{fwd}}$, the
backward-to-forward ratio $\beta$, and the per-gradient speedup
over FD for every variant.
$T_3$: plot both panels of Figure~5.
Dependencies: $T_1 \rightarrow T_2 \rightarrow T_3$.

\artout

Per-variant timing rows in \path{scaling_breakdown.csv}.
\path{plot_scaling_study.py} regenerates Figure~5.


% ---- A8: Framework decomposition ----
\newartifact

\artrel

$A_8$ is the framework-decomposition benchmark
(\path{experiments/fig4_framework_decomposition/}). It
disentangles algorithmic gains from framework overhead by timing
both ADELIA and an FD baseline under matched framework conditions
on all ten benchmark models, supporting contribution $C_4$ and
producing Figure~6.

\artexp

The observed per-gradient speedup should decompose into an
algorithmic component (reduced work from the structure-exploiting
backward pass) and a framework component (JAX/XLA vs the CuPy-based
FD baseline), with the algorithmic component accounting for a
sizeable fraction of the total.

\arttime

$\sim$3~hours on 1--4 GPUs across the ten models.

\artin

\artinpart{Hardware, Software, Installation, Inputs}

Same as $A_6$. Uses the same 10-model dataset as $A_3$ and $A_6$.

\artcomp

The workflow runs on all ten benchmark models at $P_{\min}$.
$T_1$: measure ADELIA's per-gradient wall-clock using its
default JAX/XLA backend (2 warmup + 10 timed evaluations).
$T_2$: measure the CuPy/Serinv FD baseline's per-forward-pass
wall-clock on the same hardware (2 warmup + 10 timed
evaluations).
$T_3$: compute the framework ratio
$r = t_{\mathrm{Serinv}}/t_{\mathrm{JAX}}$ and the
algorithmic speedup at parity $(2d{+}1)/(1+\beta)$,
where $\beta$ is the backward-to-forward ratio measured in
$T_1$.
$T_4$: decompose the observed per-gradient speedup into
algorithmic and framework contributions for Figure~6.
Dependencies: $T_1, T_2 \rightarrow T_3 \rightarrow T_4$.

\artout

Per-model ratios and timings in
\path{single_eval_comparison.csv},
\path{distributed_single_eval.csv}, and
\path{memory_breakdown.csv}.
\path{plot_framework_decomposition.py} regenerates Figure~6.


% ---- A9: Resource / energy efficiency ----
\newartifact

\artrel

$A_9$ is the resource and energy efficiency benchmark
(\path{experiments/fig6_resource_efficiency/}). It measures
how many GPUs FD must use (and how much energy it must draw) to
match ADELIA's 4-GPU per-gradient wall-clock on the four
production-scale multivariate models, supporting contributions
$C_3$ and $C_4$ and producing Figure~7.

\artexp

As FD is scaled up, per-gradient time ratios should decrease
while energy ratios stay flat or grow, reproducing
Figure~7. The best case for FD in the paper is WA1 at
128~GPUs, where FD is $\sim 4.2\times$ faster than ADELIA's
4-GPU run but still consumes $7.3\times$ more energy.

\arttime

$\sim$12~hours on 4--128 GPUs (ADELIA baseline at fixed 4 GPUs,
FD scaled through 4, 8, 16, 32, 64, 128 GPUs).

\artin

\artinpart{Hardware}

Same as $A_3$. The FD scaling baseline requires up to 128 GPUs
for the largest runs.

\artinpart{Software, Installation}

Same as $A_3$, additionally using the Cray Power Management
counters exposed through
\path{experiments/common/energy_monitor.py} to record energy
per gradient.

\artinpart{Datasets / Inputs}

Uses the four production-scale multivariate models SA1, WA1,
WA2, AP1 already downloaded for $A_6$.

\artcomp

The workflow runs on the four production-scale models (SA1,
WA1, AP1, WA2), all at $d{=}15$ ($2d{+}1{=}31$ FD evaluations).
$T_1$: measure ADELIA's per-gradient wall-clock and energy at
the minimum GPU count $P_{\min}$ that fits each model
(2 warmup + 10 timed evaluations).
$T_2$: for each $P \in \{4, 8, 16, 32, 64, 128\}$, run FD's
$2d{+}1$ evaluations in parallel as
$\lfloor P/P_{\min} \rfloor$ concurrent groups of $P_{\min}$
GPUs, and measure both wall-clock and total node-level energy.
$T_3$: compute the time and energy ratios of FD over ADELIA
across the scaling curve and produce both panels of Figure~7.
Dependencies: $T_1, T_2 \rightarrow T_3$.
Energy is measured via Cray PM hardware counters at the node
power-supply level (10\,Hz, covering GPU, CPU, memory, and
interconnect), with MPI barriers delimiting the measurement
window so it includes the full allocated time.

\artout

Time and energy ratio rows in
\path{resource_efficiency_all.csv}.
\path{plot_resource_energy_efficiency.py} regenerates
Figure~7.


% ---- A10: Per-stage distributed performance analysis ----
\newartifact

\artrel

$A_{10}$ is the per-stage distributed performance-analysis
benchmark
(\path{experiments/fig7_performance_analysis/}). It breaks
down the per-gradient wall-clock of the distributed two-phase
path into forward, solve, reduced-system, and gradient-prior
stages on WA1, AP1, and WA2. It supports $C_3$ and $C_4$ and
produces Figure~8.

\artexp

Per-stage breakdowns should show the relative cost of the
forward, backward solve, reduced-system aggregation, and
gradient-prior stages on the three distributed models,
reproducing Figure~8.

\arttime

$\sim$1~hour on 4 GPUs (three models).

\artin

\artinpart{Hardware, Software, Installation}

Same as $A_3$.

\artinpart{Datasets / Inputs}

WA1, AP1, WA2 inputs already downloaded for $A_6$.

\artcomp

The workflow runs on the three distributed models (WA1, AP1,
WA2) on 4 GPUs.
$T_1$: run one ADELIA gradient evaluation per model with the
distributed two-phase path instrumented to record per-stage
wall-clock for the forward pass, the backward solve, the
reduced-system aggregation, and the gradient-prior stage
(2 warmup + 10 timed evaluations, 95\% CI).
$T_2$: aggregate the per-stage contributions into the stacked
bar chart of Figure~8.
Dependencies: $T_1 \rightarrow T_2$.
Per-stage timings use synchronous CUDA event timestamps so
pipelined work is not double-counted.

\artout

Per-stage per-model rows in
\path{distributed_breakdown.csv}.
\path{plot_performance_analysis.py} regenerates Figure~8.

\end{document}
